Scalar and Representative Observables for Polarimetric SAR Data

### 2013-10-12 paper_dev/article_pp_ivm.tex
updated change ratio plots, before merging with Ian's changes

\begin{abstract}
This paper presents several novel scalar statistical models to represent the determinant of both partial and full polarimetric SAR (POLSAR) covariance matrices.
Compared to other scalar statistical models for POLSAR,
  the proposed models are highly representative of the multi-dimensional data
  and enable useful discrimination measures to be easily determined. %consequently they lead to useful discrimination measures.
The representative power of these models is explained  
when these multi-dimensional models are collapsed to single dimensional versions,
  they are shown to exactly match the traditional statistical models for the representative observable of SAR intensity.
The paper also validates the theoretical models together with their derived discrimination measures using real world POLSAR data.  
%IVMThe proposed generic models are envisioned to enable the adaptation of many existing SAR data processing techniques for use with POLSAR data.
One major benefit of this paper is that these proposed generic models can facilitate the adaptation of existing SAR techniques to higher-dimensional POLSAR data.
To illustrate, an example is given where the determinant-ratio and its additive version, contrast, are applied in the evaluation of POLSAR speckle filters.
\end{abstract}

@@

During the past decades, exponential growth in computing power has allowed once computationally-demanding Synthetic Aperture Radar (SAR)
technology to become a feasible and preferred technique for earth observation applications.
SAR technology has since extended in a few directions, one of which is polarimetric SAR (POLSAR).
POLSAR extends SAR by exploiting the natural polarization property of Electro-Magnetic (EM) waves,
  leading to the availability of multi-channel POLSAR data, compared to traditional one-channel SAR data.

While statistical models are important in understanding the stochastic nature of both SAR and POLSAR, in extending our understanding towards multi-dimensional POLSAR, an important issue needs to be addressed.
That is there now exists not one but many observable quantities in multi-channel POLSAR (as opposed to single-channel SAR intensity).
The research community has responded largely by developing different statistical models for various POLSAR observables.
However for a statistical model to be useful, scalar discrimination measures generally need to be derivable from it.
Thus, practical applications for POLSAR data processing require dis-similarity measures that are scalar, consistent and preferably homoskedastic on the one hand.
On the other hand, the observable quantity being modelled should be naturally representative of the high dimensionality POLSAR data.

This paper presents novel statistical models for multi-channel POLSAR data as generalisations of the commonly used and representative statistical models for the one-channel SAR data.
This approach makes feasible the adaptation of many existing SAR processing techniques to be applied to POLSAR data.
After section \ref{sec:lit_review} reviews related work in published literature, statistical models for the heteroskedastic POLSAR determinant and its homoskedastic log-domain equivalent, together with several other discrimination measures are derived in section  \ref{sec:theoretical_model}.
The representative properties of these observables are then demonstrated in section \ref{sec:sar_special_case_of_polsar}
  where the multi-dimensional POLSAR data is collapsed to a one-dimensional SAR scenario.
With this transformation, the determinant of the POLSAR covariance matrix is shown to be equivalent to SAR intensity,
%  which also brings many standard statistical models used with SAR  under the umbrella of these generic POLAR models.
  and allow many conventional SAR statistical models to stay under the umbrella of these generic POLAR models.
The proposed theoretical models are then validated against real life practical data in section \ref{sec:polsar_models_validation}.
Section \ref{sec:discussion} briefly describes how evaluation of an existing SAR data processing technique, the speckle filter, can be extended to  POLSAR data processing using this approach.
This is followed by a conclusion in Section \ref{sec:conclusion}.

@@

This section outlines related published work.
In particular, section \ref{sec:lit_models} reviews various scalar statistical models used for different POLSAR observables.
Its purpose is to show that none of those  models have led to statistically consistent discrimination measures. 
Section \ref{sec:lit_measures} further strengthens those findings by discussing discrimination measures that have been proposed for POLSAR,
  and demonstrates that almost all of them are based on the likelihood statistical test for complex Wishart distribution.
However, while an exact statistical distribution is needed for the test,
  only asymptotic distribution is used in the original work \cite{Conradsen_2003_TGRS_4}.

@@

The commonly used measure of distance for matrices are either the Euclidean or Manhattan distances, defined as:
\begin{align}
  d(C_x,C_y) &= \sum_{i,j} |\mathbb{R} (C_x - C_y)_{i,j}| + \sum_{i,j} |\mathbb{I} (C_x - C_y)_{i,j}| \\
  d(C_x,C_y) &= \sqrt{\sum_{i,j} |C_x - C_y|_{i,j}^2 }
\end{align}
where $C_{i,j}$ denotes the (i,j) elements of the POLSAR covariance matrix C,
 $||$ denotes absolute values
and $\mathbb{R},\mathbb{I}$ denote the real and imaginary parts respectively.
However, in the context of POLSAR% covariance matrix
, these dis-similarity measures are not widely used 
  mainly due to the multiplicative nature of the noisy data.

@@

Recent works have suggested alternative dissimilarity measures including the symmetric and asymmetric refined Wishart distance \cite{Anfinsen_2007_ESA_POLINSAR},
\begin{align}
  d(C_x,C_y) &= \frac{1}{2} tr(C_x^{-1}C_y + C_y^{-1}C_x) - d \\
    d(C_x,C_y) &= \ln|C_x| - \ln|C_y| + tr(C_xC_y^{-1}) - d
\end{align}
the Bartlett distance \cite{Kersten_2005_TGRS_519},
  \begin{align}
  d(C_x,C_y) &= 2 \ln |C_{x+y}| - \ln |C_x| - \ln |C_y| - 2d\ln2
  \end{align}
the Bhattacharyya distance \cite{Lee_2011_IGARSS_3740},
\begin{equation}
  r(C_x,C_y) = \frac{|C_x|^{1/2} |C_y|^{1/2}}{|(C_x+C_y)/2|}
\end{equation}
and the Wishart Statistical test distance \cite{Cao_2007_TGRS_3454}
\begin{equation}
  d(C_x,C_y) = (L_x + L_y) \ln|C| - L_x \ln|C_x| - L_y\ln|C_y|
\end{equation}
.

%%***IVM: Hai, you haven't introduced the notation for these!! What is tr() (in eqn. 3), what is Lx, Ly.... I changed your A_{i,j} to C_{i,j} above, just after eqn.2

Closer investigation of these dis-similarity measures reveals that most are related in some ways. %to each other.
The Bhattacharyya distance is easily shown to be related to the Barlett distance.
At the same time the Barlett distance can be considered a special case of the Wishart Statistical Test distance,
  when the two data sets have the same number of looks, i.e. $L_x=L_y$.
The close relation among the measures is further supported by the fact that
  all of their proposing papers refered the same statistical model developed in \cite{Conradsen_2003_TGRS_4} as a foundation.
In \cite{Conradsen_2003_TGRS_4}, to determine if the two scaled multi-look POLSAR covariance matrix $Z_x$ and $Z_y$,
  which have $L_x$ and $L_y$ as the corresponding number of looks,
  come from the same underlying stochastic process,
the likelihood ratio statistics for POLSAR covariance matrix is considered:  
\begin{equation*}
  Q = \frac{(L_x+L_y)^{d \cdot (L_x+L_y)}}{L_x^{d \cdot L_x} L_y^{d \cdot L_y}} \frac{|Z_x|^{L_x} |Z_y|^{L_y} }{|Z_x+Z_y|^{(L_x+L_y)}}
\end{equation*}

Taking the log-transformation of the above equation, and note that $C_{vx} = Z_x / L_x$, $C_{vy} = Z_y / L_y$ and $C_{vxy} = (Z_x + Z_y)/(L_x + L_y)$ then:
\begin{eqnarray}
  Q &=& \frac{|C_{vx}|^{L_x} \cdot |C_{vy}|^{L_y} }{|C_{vxy}|^{L_x + L_y}} \label{eqn:ori_likelyhood_stats} \\
  \ln Q &=& L_x \ln |C_{vx}| + L_y \ln |C_{vy}| - (L_x + L_y) \ln |C_{vxy}| \label{eqn:log_likelyhood_stats}
\end{eqnarray}

To detect changes, a test statistic is developed based on this measure of distance.
This means a distribution is to be derived for the dissimilarity measure.
However, in the original work \cite{Conradsen_2003_TGRS_4}, only an asymptotic distribution has been proposed.

\section{Scalar Statistical Models for POLSAR}  
\label{sec:theoretical_model}

In this section, several scalar statistical models for POLSAR are presented.
Section \ref{sec:model_basic} briefly introduces the basic foundations, concepts and notation used in this paper, %reviews related work in published literature,
  while section \ref{sec:polsar_heterosked_model_and_log_transform}
  presents the theoretical models together with mathematical evidence for two related points.
The first point is that the POLSAR data is multiplicative and heteroskedastic in its original domain.
And the second conclusion states that log-transformation converts it into an additive and homoskedastic model.
Some consistent measures of distance are then derived and presented in section \ref{sec:distance_measure}.

\subsection{Basics of POLSAR Statistical Analysis}
\label{sec:model_basic}

In this paper, the POLSAR scattering vector is denoted as $s$.
In the case of partial polarimetric SAR (single polarization in transmit and dual polarization in receipt),
  the vector is two-dimensional ($d=2$) and is normally written as: 
\begin{equation}
s_{part}=\begin{bmatrix}
S_h\\ 
S_v
\end{bmatrix}
\end{equation}
In the case of full and monostatic POLSAR data,
  the vector is three-dimensional ($d=3$) and is presented as:
\begin{equation}
s_{full}=\begin{bmatrix}
S_{hh}\\
\sqrt{2}S_{hv}\\
S_{vv}
\end{bmatrix}
\end{equation}

@@

The approach taken in this paper differs by applying the homoskedastic log transformation  on a less-than-well-known result for the determinant of the covariance matrix.
Goodman \cite{Goodman_1963_AMS_178} proved
that the ratio between the observable and expected values of the sample covariance matrix determinants
  behave like a product of $d$ chi-squared random variables with different degrees of freedom: 
\begin{equation}
\chi^d_L = (2L)^d \frac{|C_v|}{|\Sigma_v|} \sim \prod_{i=0}^{d-1} \chi^2 (2L-2i)
\label{eqn:prod_chi_squared_rv}  
\end{equation}

Its log-transformed variable consequently 
  behaves like a summation of $d$ log-chi-squared random variables with the same degrees of freedom  
\begin{equation}
\Lambda^d_L = ln \left[ (2L)^d \frac{|C_v|}{|\Sigma_v|} \right] \sim \sum_{i=0}^{d-1} \Lambda^\chi (2L-2i)
\label{eqn:sum_log_chi_squared_rv}
\end{equation}
with
  $\Lambda^\chi (k) \sim \ln \left[ \chi^2 (k) \right]$.
We will now use this result to develop the statistical models for POLSAR. %log-transformed measures of distance.
%***IVM added the sentence above... **Please check it's true!!!!

\subsection{Original Hetoroskedastic Domain and the Homoskedastic Log-Transformation}
\label{sec:polsar_heterosked_model_and_log_transform}

In this section the multiplicative nature of POLSAR data is first illustrated.
 Log-transformation is then applied to convert  the data into a more familiar additive model.
Heteroskedasticity, which is defined as the dependence of variance upon the underlying signal,
  is shown to be present in the original POLSAR data.
In the log-transformed domain, a homoskedastic model is demonstrated,
  where sample variance is fixed and thus independent of the underlying signal.
Several intermediate mathematical steps are not shown here, but can be found in Appendix \ref{chap:appendix_a}.

From Eqns. \ref{eqn:prod_chi_squared_rv} and \ref{eqn:sum_log_chi_squared_rv}
we can deduce the following relationships:
\begin{eqnarray}
  |C_v| &\sim& |\Sigma_v| \cdot \frac{1}{(2L)^d} \cdot \prod_{i=0}^{d-1} \chi^2 (2L-2i) \label{eqn:determinant_distribution} \\
  \ln|C_v| &\sim& \ln|\Sigma_v| - d \cdot \ln(2L) + \sum^{d-1}_{i=0} \Lambda(2L-2i)
\label{eqn:log_determinant_distribution}  
\end{eqnarray}

In a given homogeneous POLSAR area, the parameters $\Sigma_v$, $d$ and $L$ can be considered constant.
Thus Eqn. \ref{eqn:determinant_distribution} indicates that 
  in the original POLSAR domain, a multiplicative speckle noise pattern is present.
At the same time, Eqn. \ref{eqn:log_determinant_distribution} shows that
  the logarithmic transformation has converted this to the more familiar additive noise.

Since chi-squared random variables $X\ \sim\ \chi^2(k)\ $ follow a known PDF:
\begin{equation}
pdf(x;2L) =
  \frac{x^{L-1} e^{-x/2}}{2^L \Gamma\left(L\right)}
\label{eqn:chi_squared_dist_pdf:chap4}
\end{equation}
applying the variable change theorem, 
  its log-transformed variable follows the PDF of:
\begin{equation}
  pdf(x;2L=k) = \frac{e^{Lx-e^x/2}}{2^{L}\Gamma(L)}
\end{equation}

Using these PDFs, the characteristic functions (CF) of both the chi-squared and log-chi-squared random variables
  can be written as:
  \begin{eqnarray}
    CF_\chi(t) &=& (1-2it)^{−L} \\ 
    CF_\Lambda(t) &=& 2^{it} \frac{\Gamma(L+it)}{\Gamma(L)} \label{eqn:log_chi_squared_characteristic_function}
  \end{eqnarray}
 Their means and variances can subsequently be computed from these  characteristic functions, leading to the following relationships:  
  \begin{eqnarray}
    avg \left[ \chi(2L) \right]&=&2L \\
var \left[ \chi(2L) \right]&=&4L \\
avg \left[ \Lambda(2L) \right] &=& \psi^0(L) + \ln2 \\
var \left[ \Lambda(2L) \right] &=& \psi^1(L)
  \end{eqnarray}
  where $\psi^0()$ and $\psi^1()$ represent the digamma and trigamma functions respectively.

Since the average and variance of both chi-squared distribution and log-chi-squared distribution are constant,
  the product and summation of these random variables also has fixed summary statistics.
Specifically:
\begin{align*}
  avg \left[ \prod^{d-1}_{i=0} \chi^2(2L-2i) \right] &= 2^d \cdot \prod^{d-1}_{i=0} (L-i), \\
  var \left[ \prod^{d-1}_{i=0} \chi^2(2L-2i) \right] &= \prod^{d-1}_{i=0} 4(L-i)(L-i+1) - \prod^{d-1}_{i=0} 4(L-i)^2, \\
  avg \left[ \sum^{d-1}_{i=0} \Lambda(2L-2i) \right] &= d \cdot \ln{2} + \sum^{d-1}_{i=0} \psi^0(L-i), \\
  var \left[ \sum^{d-1}_{i=0} \Lambda(2L-2i) \right] &= \sum^{d-1}_{i=0} \psi^1(L-i)
\end{align*}

Combining these results with Eqns. \ref{eqn:determinant_distribution} and \ref{eqn:log_determinant_distribution}, we have:
\begin{align}
  avg \left[ |C_v| \right]  &= \frac{|\Sigma_v|}{L^d} \prod^{d-1}_{i=0} (L-i)\\
  var \left[ |C_v| \right]  &=   \frac{|\Sigma_v|^2 \left[ \prod^{d-1}_{i=0} (L-i)(L-i+1) - \prod^{d-1}_{i=0} (L-i)^2 \right] }{L^{2d}} \label{eqn:var_det_is_heteroskedastic}\\
  avg \left[ \ln |C_v| \right] &= \ln |\Sigma_v| - d \cdot \ln{L}  + \sum^{d-1}_{i=0} \psi^0(L-i) \label{eqn:avg_log_det} \\
  var \left[ \ln |C_v| \right] &=  \sum^{d-1}_{i=0} \psi^1(L-i) \label{eqn:var_log_det_is_homoskedastic}
\end{align}

>>>>

\begin{abstract}
%Problem statement  
This paper proposes a scalar and representative observable for multi-dimensional POLSAR data,
  from which statistically consistent discrimination measures can be derived.
%What actually were done? and As the result, what is learned?
Specifically, the statistical behaviour of the POLSAR covariance matrix determinant is used
  to derive a scalar and generic statistical model for multi-dimensional POLSAR data,
  which is specifically applicable to the two and three dimensional versions of partial and full monostatic polarimetric SAR data.
As the POLSAR covariance matrix determinant generalizes the SAR intensity towards the multidimensional extension,
  the proposed model is also shown to include the traditional model of SAR intensity under its natural coverage.
%What are the larger implications of the findings presented in this paper?
Consequently, the main beneficial implication of the proposed approach is that
  it provides a consistent theory unifying the currently disconnected proposals for SAR and POLSAR discrimination measures
  which simplifies the adaptation of existing SAR data processing techniques for POLSAR data.
%  it enables the adaptation of many existing SAR data processing techniques for POLSAR data,
%  and at the same time, it provides a consistent theory unifying the seemingly disparate discrimination measure proposals for SAR and POLSAR data.
\end{abstract}

@@

%What is the situation, the background context of your research? the BIG problem!
During the past decades, exponential growth in computing power has allowed the once computationally-demanding Synthetic Aperture Radar (SAR)
technology to become a feasible and preferred technique for earth observation applications.
SAR technology has since been extended in a few directions, one of which is polarimetric SAR (POLSAR).
POLSAR extends SAR by exploiting the natural polarization property of Electro-Magnetic (EM) waves,
  leading to the availability of multi-channel POLSAR data, compared to traditional one-channel SAR data.

%Research Motivation: Why it is important to solve this problem?
%The POLSAR data, however, is multidimensional and stochastic.
Like SAR data, POLSAR data is stochastic.
Moreover, it is multi-dimensional, making it even harder to interpret.
Under this context, it is therefore important to establish a simple and intuitive understanding of the data.
 Statistical models are undoubtedly crucial in understanding its stochastic nature.
While several %multidimensional 
models have been proposed to describe POLSAR data, they
  unfortunately tend to be complex and not very intuitive due to the multidimensional nature of the data.
Practical POLSAR data processing however, makes heavy use of scalar discrimination measures,
  which should be based on statistically consistent models for scalar and representative observables of the multidimensional data.
It is thus important to establish a scalar and representative observable for this multi-dimensional POLSAR data.

%Moreover, to provide theoretical foundation for discrimination measures proposal,
%  statistically consistent models should be derived from this scalar observable.

%How do others solve this problem?
%The complex POLSAR data have been statistically modelled as following the complex Wishart distribution,
%  which apparently is multidimensional, complex and thus not very intuitive.
There has been a few scalar observables with accompanying statistical models proposal for POLSAR \cite{Conradsen_2003_TGRS_4, Alberga_2008_IJRS_4129, Joughin_1994_TGRS_562, Lee_1994_TGRS_1017, Touzi_1996_TGRS_519, Lopez-Martinez_2003_TGRS_2232, Erten_2012_Sensors_2766},
  but none of them is able to provide meaningful scalar discrimination measures.
%  but none of them has led to scalar discrimination measures proposal.
As such no observable has been widely accepted as being highly representative of this multi-dimensional data,
  which severely limits their applicability in practical data processing applications.
From an alternative approach, a few POLSAR discrimination measures have been proposed,
  but all are based on the likelihood ratio statistics.
This statistical test should be based on an exact and consistent statistical distribution
  but so far has only shown to be based on an asymptotic distribution.

%Research Objective: Give your precise problem statement  
This article hence presents a scalar and representative observable, and its associated generic statistical model to describe multi-dimensional POLSAR data and provide a consistent foundation for the derivation of POLSAR discrimination measures.
%Now State your approach  
%This study started off with the observation that the determinant of the covariance matrix is widely used in POLSAR discrimination measures.
%However, our survey of existing scalar models for POLSAR indicates that no statistical model for this observable has been published so far.
%Consequently, in this article, the statistical behaviour of the determinant of POLSAR covariance matrix is further investigated.
%Prof Vuns: comment: the article provides the result, not the journey!
POLSAR data have been statistically modelled as following the complex Wishart distribution.
Consequently, the generic statistical model for the covariance matrix determinant is presented as being just a scalar projection of the multidimensional data.
This model is then used to derive several scalar and consistent statistical descriptions suggesting that their associated observables are capable of being used as discrimination measures for POLSAR data.
%
%***IVM: the following sentence doesn't make sense, also, I don't think it's good to talk about 'models of a model'.  Maybe 'model variants' instead?
%%%***IVM***The specific two and three dimensional models of these models, which %also include the derived discrimination measures, are originally designed against both partial and full polarimetric SAR data.

%The specific two and three dimensional models are then validated against both partial and full polarimetric SAR data.
Interestingly the paper also shows that %It is also shown that
  the specific one dimensional versions of the proposed  models 
  match perfectly with the traditional statistical model used for SAR intensity.
This effectively incorporates the common one dimensional SAR theory under the unbrella of the proposed scalar approach for multi dimensional POLSAR.
With this new insight the different discrimination proposals for SAR and POLSAR are reviewed and it is shown that the approach proposed in this paper provides a strong, unifying and consistent foundation relating them all together.
The applicability of these theoretical models will be illustrated by experiments where the specific one, two and three dimensional versions of the proposed generic models are validated against practical data. 
%Furthermore, the approach proposed in this paper
%  not only provides a consistent foundation for the different discrimination measures proposal in POLSAR
%  but also leads to the establishment of a new discriminating measures: the determinant-ratio and the change-ratio.

%What is the novelty of your solution? What are its advantages
%Compared to other published scalar observables with accompanying statistical model,
%  the proposed observable and its models are highly representative of the multidimensional data.
%The representative power of the proposed observable is justified by the following reasons.
%Firstly, the covariance matrix determinant, when collapsed into one-dimensional case, transforms neatly into the representative SAR intensity.
%Secondly, the statistical model for this observable is shown as the generic multi-dimensional extension of the traditional and representative intensity model for the one dimensional SAR data.
%And last but not least, this observable also leads to the determinant-ratio discrimination measure which
%  not only is simpler in concept and computation than the existing POLSAR discrimination measures
%  but also can be considered as the natural multi-dimensional extension of the commonly-used intensity-ratio discrimination measure in SAR.
% Prof Vun Comment: already mentioned above, no need to repeat
%hai thoughts: actually this should be included to justify our title, and our approach. It is however being moved to discussion / conclusion section

%How do you know if the problem is solved? State your research contribution to be achieved
%In summary, the main objective of this article is to propose the determinant of the POLSAR covariance matrix as a scalar observable
%  that is representative of the multi-dimensional data.
%Specifically the following research results are to be achieved:
%  1. The generic scalar and representative statistical model for the determinant of the multi-dimensional POLSAR covariance matrix is to be derived, 
%  2. This generic model is also shown leading to staititically consistent models for newly proposed POLSAR observables, namely the determinant-ratio and the change-ratio.
%  3. The special one-dimensional version of the derived generic model is to be shown perfectly matches with the traditionally used SAR intensity model,
%  4. All specific one, two and three dimensional versions of the derived generic model are to be validated against real-life captured data,
%(Hai NOTE: I kept this to keep my focus while working on this paper.
%  It can be removed before the final submission.)

%Article Organization: Show that you deliver on the objectives and hence solve the stated problem
The remainder of this article is organized as follows.
After the next section reviews existing discrimination measures as well as scalar observables models for POLSAR, section \ref{sec:theoretical_model}
  derives the generic statistical model for the determinant of the POLSAR covariance matrix
  and proposes several new discrimination measures for the multi-dimensional data.
After section \ref{sec:sar_special_case_of_polsar} illustrates the match between the specific model of one dimensional and the traditional model for SAR intensity,
  section \ref{sec:link_sar_polsar} links the disparate discrimination measure proposals for the traditional SAR and the more recent POLSAR data.
%Section \ref{sec:polsar_models_validation} then validates the specific one, two and three dimensional versions of the proposed model against real-life captured data.
The applicability of these models against practical data is illustrated in section \ref{sec:polsar_models_validation}.
Section \ref{sec:discussion} presents a high-level discussion of the results presented before
  section \ref{sec:conclusion} finally concludes the paper.

@@

%This section outlines related publication.
Section \ref{sec:lit_models} reviews various statistical models used for different scalar observables for POLSAR.
  where it is shown that none are able to provide statistically consistent discrimination measures. 
Section \ref{sec:lit_measures} further strengthens those findings by discussing discrimination measures that have been proposed for POLSAR,
  and demonstrates that almost all of them are based on the likelihood statistical test for complex Wishart distribution.
While an exact statistical distribution is expected to be necessary for a test,
  only an asymptotic distribution is used in the underlying approach \cite{Conradsen_2003_TGRS_4}.

@@

The commonly used measure of distance for matrices are either the Euclidean or Manhattan distances, defined respectively as:
\begin{align}
  d(C_x,C_y) &= \sum_{i,j} |\mathbb{R} (C_x - C_y)_{i,j}| + \sum_{i,j} |\mathbb{I} (C_x - C_y)_{i,j}| \\
  d(C_x,C_y) &= \sqrt{\sum_{i,j} |C_x - C_y|_{i,j}^2 }
\end{align}
where $C_{i,j}$ denotes the $(i,j)$ elements of the POLSAR covariance matrix C,
 $||$ denotes absolute values
and $\mathbb{R},\mathbb{I}$ denote the real and imaginary parts respectively.
However, in the context of POLSAR, % covariance matrix
  these dissimilarity measures are not widely used 
  mainly due to the multiplicative nature of the noisy data.

@@

Recent works have suggested alternative dissimilarity measures including the symmetric and asymmetric refined Wishart distance \cite{Anfinsen_2007_ESA_POLINSAR},
\begin{align}
  d(C_x,C_y) &= \frac{1}{2} tr(C_x^{-1}C_y + C_y^{-1}C_x) - d \\
    d(C_x,C_y) &= \ln|C_x| - \ln|C_y| + tr(C_xC_y^{-1}) - d
\end{align}
the Bartlett distance \cite{Kersten_2005_TGRS_519},
  \begin{align}
  d(C_x,C_y) &= 2 \ln |C_{x+y}| - \ln |C_x| - \ln |C_y| - 2d\ln2
  \end{align}
the Bhattacharyya distance \cite{Lee_2011_IGARSS_3740},
\begin{equation}
  r(C_x,C_y) = \frac{|C_x|^{1/2} |C_y|^{1/2}}{|(C_x+C_y)/2|}
\end{equation}
and the Wishart Statistical test distance \cite{Cao_2007_TGRS_3454}
\begin{equation}
  d(C_x,C_y) = (L_x + L_y) \ln|C| - L_x \ln|C_x| - L_y\ln|C_y|
\end{equation}

Closer examination of these dissimilarity measures reveals that most of them are related in some ways. %to each other.
The Bhattacharyya distance is easily shown to be related to the Barlett distance.
At the same time the Barlett distance can be considered a special case of the Wishart Statistical Test distance,
  when the two data sets have the same number of looks, i.e. $L_x=L_y$.
The close relation among the measures is further supported by the fact that
  all of their publications refered the same statistical model developed in \cite{Conradsen_2003_TGRS_4} as the foundation.
In \cite{Conradsen_2003_TGRS_4}, to determine if the two scaled multi-look POLSAR covariance matrixes $Z_x$ and $Z_y$,
  which have $L_x$ and $L_y$ as the corresponding number of looks,
  come from the same underlying stochastic process,
the likelihood ratio statistics for POLSAR covariance matrix is considered:  
\begin{equation*}
  Q = \frac{(L_x+L_y)^{d \cdot (L_x+L_y)}}{L_x^{d \cdot L_x} L_y^{d \cdot L_y}} \frac{|Z_x|^{L_x} |Z_y|^{L_y} }{|Z_x+Z_y|^{(L_x+L_y)}}
\end{equation*}

Taking the log-transformation of the above equation, and denoting $C_{vx} = Z_x / L_x$, $C_{vy} = Z_y / L_y$ and $C_{vxy} = (Z_x + Z_y)/(L_x + L_y)$ then:
{\small
\begin{align}
  Q &= \frac{|C_{vx}|^{L_x} \cdot |C_{vy}|^{L_y} }{|C_{vxy}|^{L_x + L_y}} \label{eqn:ori_likelyhood_stats} \\
  \ln Q &= L_x \ln |C_{vx}| + L_y \ln |C_{vy}| - (L_x + L_y) \ln |C_{vxy}| \label{eqn:log_likelyhood_stats}
\end{align}
}

To detect changes, a test statistic is developed for this discrimination measure.
This means a distribution is to be derived for the dissimilarity measure.
However, in the original work \cite{Conradsen_2003_TGRS_4}, only an asymptotic distribution was used.
This paper proposes a statistical model for the determinant of the POLSAR covariance matrix $|C_v|$
  which is capable of providing an exact distribution for the test.

\section{The Generic Scalar Statistical Model for POLSAR}  
\label{sec:theoretical_model}

%In this section, the generic scalar statistical model for POLSAR is presented.
In this section, after the basic foundations of POLSAR statistical analysis are introduced,
  the generic scalar statistical model for the multi-dimensional POLSAR will be presented.
%  this section presents the theoretical model for multi-dimensional POLSAR data.

In this paper, the POLSAR scattering vector is denoted as $s$.
For partial polarimetric SAR (single polarization in transmit and dual polarization in receipt),
  the vector is two-dimensional ($d=2$) and is normally written as: 
\begin{equation}
s_{part}=\begin{bmatrix}
S_h\\ 
S_v
\end{bmatrix}
\end{equation}
For full and monostatic POLSAR data,
  the vector is three-dimensional ($d=3$) and is presented as:
\begin{equation}
s_{full}=\begin{bmatrix}
S_{hh}\\
\sqrt{2}S_{hv}\\
S_{vv}
\end{bmatrix}
\end{equation}

@@

The approach taken in this paper differs by applying the homoskedastic log transformation  on a less-than-well-known result for the determinant of the covariance matrix.
Goodman \cite{Goodman_1963_AMS_178} proved
that the ratio between the observable and expected values of the sample covariance matrix determinants
  behaves like a product of $d$ chi-squared random variables with different degrees of freedom: 
\begin{equation}
\chi^d_L = (2L)^d \frac{|C_v|}{|\Sigma_v|} \sim \prod_{i=0}^{d-1} \chi (2L-2i)
\label{eqn:prod_chi_squared_rv}  
\end{equation}

%We will now use 
This result is used here to develop the generic scalar statistical model for POLSAR. %log-transformed measures of distance.
From Eqn. \ref{eqn:prod_chi_squared_rv} %and \ref{eqn:sum_log_chi_squared_rv}
we have: %can deduce the following relationships:
\begin{eqnarray}
  |C_v| &\sim& |\Sigma_v| \cdot \frac{1}{(2L)^d} \cdot \prod_{i=0}^{d-1} \chi (2L-2i) \label{eqn:determinant_distribution} %\\
%  \ln|C_v| &\sim& \ln|\Sigma_v| - d \cdot \ln(2L) + \sum^{d-1}_{i=0} \Lambda(2L-2i)
%\label{eqn:log_determinant_distribution}  
\end{eqnarray}

In a given homogeneous POLSAR area, the parameters $\Sigma_v$, $d$ and $L$ can be considered constant.
Thus Eqn. \ref{eqn:determinant_distribution} indicates that 
  in the original POLSAR domain, a multiplicative speckle noise pattern is present.

Moreover, since the average and variance of these chi-squared distributions %and log-chi-squared distribution
  are known to be constant, i.e. $avg \left[ \chi(2L) \right] = 2L$ and $var \left[ \chi(2L) \right] = 4L$,
  the product and summation of these random variables also have fixed summary statistics.
Specifically:
{\footnotesize
\begin{align*}
  avg \left[ \prod^{d-1}_{i=0} \chi(2L-2i) \right] &= 2^d \cdot \prod^{d-1}_{i=0} (L-i), \\
  var \left[ \prod^{d-1}_{i=0} \chi(2L-2i) \right] &= \prod^{d-1}_{i=0} 4(L-i)(L-i+1) - \prod^{d-1}_{i=0} 4(L-i)^2, %\\
%  avg \left[ \sum^{d-1}_{i=0} \Lambda(2L-2i) \right] &= d \cdot \ln{2} + \sum^{d-1}_{i=0} \psi^0(L-i), \\
%  var \left[ \sum^{d-1}_{i=0} \Lambda(2L-2i) \right] &= \sum^{d-1}_{i=0} \psi^1(L-i)
\end{align*}
}

Combining these results with Eqn. \ref{eqn:determinant_distribution} %and \ref{eqn:log_determinant_distribution}
  , we have:
{\footnotesize
\begin{align}
  avg \left[ |C_v| \right]  &= \frac{|\Sigma_v|}{L^d} \prod^{d-1}_{i=0} (L-i)\\
  var \left[ |C_v| \right]  &=   \frac{|\Sigma_v|^2 \left[ \prod^{d-1}_{i=0} (L-i)(L-i+1) - \prod^{d-1}_{i=0} (L-i)^2 \right] }{L^{2d}} \label{eqn:var_det_is_heteroskedastic}%\\
%  avg \left[ \ln |C_v| \right] &= \ln |\Sigma_v| - d \cdot \ln{L}  + \sum^{d-1}_{i=0} \psi^0(L-i) \label{eqn:avg_log_det} \\
%  var \left[ \ln |C_v| \right] &=  \sum^{d-1}_{i=0} \psi^1(L-i) \label{eqn:var_log_det_is_homoskedastic}
\end{align}
}

For a real world captured image, while the parameters $d$ and $L$ do not change for the whole image,
  the underlying $\Sigma_v$ is expected to differ from one region to the next.
Thus over a heterogeneous scene, the stochastic process for $|C_v|$ and $\ln |C_v|$ vary depending on the underlying signal $\Sigma_v$. 
In such context, Eqn. \ref{eqn:var_det_is_heteroskedastic} implies that the variance of $|C_v|$ also differs depending on the underlying signal $\Sigma_v$ (i.e. it is   heteroskedastic).

Similar to the way intensity-ratio is proposed as the discrimination measure for the multiplicative and heteroskedastic SAR intensity \cite{Rignot_1993_TGRS_896},
in this paper, the determinant-ratio and the change-ratio are also proposed as a discrimination measure for the POLSAR data.
  which is shown above to also suffer from the multiplicative and heteroskedastic phenomena.
%Similar to the way dispersion and contrast is defined for one-dimensional SAR \cite{Le_2010_ACRS},
%  this section introduces the consistent sense of distance for POLSAR.

%Assuming, on the one hand,
For cases where the true value of the underlying signal $\Sigma_v$ is known \textit{a priori},
then the determinant-ratio of the signal random variable ($\mathbb{R}_{\Sigma}$) %and log-distance ($\mathbb{L}$)
%  are observable according to the following definitions:
  is defined as:
%Eqns. \ref{eqn:prod_chi_squared_rv} and \ref{eqn:sum_log_chi_squared_rv} lead straight to the definition of the following random variables, which is the :
\begin{eqnarray}
  \mathbb{R}_{\Sigma} &=& \frac{|C_v|}{|\Sigma_v|} \label{eqn:determinant_ratio_observables}%\\
%  \mathbb{L} &=& \ln|C_v| - \ln|\Sigma_v| \label{eqn:log_distance_observables} 
\end{eqnarray}
%On the other hand, under a looser assumption %From another perspective
For cases where the POLSAR data is known to have come from a homogeneous area, but the true value of the underlying signal $\Sigma_v$ is \textit{unknown},
 then a random variable called the change-ratio ($\mathbb{R}_{C}$)  is defined as:
\begin{eqnarray}
  \mathbb{R}_{C} &=& \frac{|C_1|}{|C_2|} \label{eqn:determinant_ratio_samples_observables}%\\
%  \mathbb{L} &=& \ln|C_v| - \ln|\Sigma_v| \label{eqn:log_distance_observables} 
\end{eqnarray}
where $C_1$ and $C_2$ are samples of the covariance matrix determinant in an assumed homogeneous area.

%  the dispersion ($\mathbb{D}$) and contrast ($\mathbb{C}$) random variables are the observables defined as:
%\begin{eqnarray}
%  \mathbb{D} &=& \ln{|C_v|} - avg(\ln{|C_v|}) \label{eqn:dispersion_observable}\\
%  \mathbb{C} &=& \ln(|C_{v1}|) - \ln(|C_{v2}|) \label{eqn:contrast_observable}
%\end{eqnarray}
